---
title: "hw1_Mauro"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
democracy <- read.csv("democracy.csv", stringsAsFactors = FALSE, na.strings = ".")
```

## Initial Data Exploration

A summary of all variables in the democracy dataset:
```{r}
summary(democracy)

```

Histogram of 'Political Liberties'

```{r}
hist(democracy$POLLIB, xlab = "Political Liberties", main = "Histogram of Political Liberties")
```

Historgram of 'GDP per capita'

```{r}
hist(democracy$GDPW, xlab = "GDP per capita", main = "Histogram of GDP per capita")
```

Histogram of the log of 'GDP per capita.' The difference is that the log of GDP per capita makes the data more evenly distributed. 

```{r}
hist(log10(democracy$GDPW), xlab = "GDP per capita(log)", main = "Histogram of GDP per capita (log)")
```

## Exploring the Relationships

Scatterplot of 'Politics Liberties' and 'GDP per capita'

```{r}
plot(democracy$POLLIB, democracy$GDPW, xlab = "Political Liberties", ylab = "GDP per capita")
ggplot(democracy, aes(x = POLLIB, y = GDPW)) + 
  geom_point() + 
  geom_jitter(width = 0) +
  labs(x = "Political Liberties", y = "GDP per capita")
```

Scatterplot of 'Political Liberties' and 'GDP per capita(log).' The logged distribution has less skew at the top end of the distribution (higher GDP per capita).

```{r}
ggplot(democracy, aes(x = POLLIB, y = log10(GDPW))) + 
  geom_point() + 
  geom_jitter(width = 0) +
  labs(x = "Political Liberties", y = "GDP per capita")
```

Boxplot of 'GDP per capita' for both oil and non-oil producing countries

```{r}
nooil <- cbind(democracy$OIL==0)
       
ggplot(democracy, aes(x=nooil, y=GDPW)) +
  geom_boxplot() +
  labs(y = "GDP per capita", x = "Oil Producing")
```

## Transforming and Analyzing Data

Mean 'GDP per capita' of Catholic countries. The mean GDP of countries with greater than 40 percent of Catholics is lower than the mean GDP of all global countries.

```{r}
mean(democracy$GDPW)

cathmean <- democracy %>% filter(CATH >= 40) %>% select(GDPW) %>% glimpse()
summary(cathmean)
```

Mean 'GDP per capita' of differing ethnolinguistic fractionalization levels among countries.

```{r}
lingomean <- democracy %>% filter(ELF60 > .60) %>% select(GDPW) %>% glimpse()
summary(lingomean)

lingomean2 <- democracy %>% filter(ELF60 < .60) %>% select(GDPW) %>% glimpse()
summary(lingomean2)

lingomean3 <- democracy %>% filter(is.na(ELF60)) %>% select(GDPW) %>% glimpse()
summary(lingomean3)
```

Median of the average years of education in all countries in 1985

```{r}
demoeducation <- democracy %>% filter(YEAR == 1985) %>% glimpse()
summary(demoeducation$EDT)
```


The closest country to the median (Venezuela)

```{r}
democracy %>% filter(EDT == 5.625 & YEAR == 1985) %>% glimpse()
```


Median of the average years of education among democracies in 1985

```{r}
demoeducation2 <- democracy %>% filter(YEAR == 1985 & REG == 1) %>% glimpse()
summary(demoeducation2$EDT)
```

The closest country to the median (Mauritius)

```{r}
democracy %>% filter(EDT == 7.805 & YEAR == 1985 & REG == 1) %>% glimpse()
```

The 25th and 75th percentiles of ethnolinguistic fractionalization in new and old countries

```{r}
frac <- democracy %>% filter(NEWC == 1) %>% glimpse() %>% select(ELF60)
summary(frac)

frac2 <- democracy %>% filter(NEWC == 0) %>% glimpse() %>% select(ELF60)
summary(frac2)
```


